Mutual Modelling

### 2016-01-25 mutual-modeling.tex
Added back Mirweis
Typos reported by reviewers

% ANON
%\author{
%    Pierre Dillenbourg
%    \and
%    Séverin Lemaignan
%    \and
%    Nicolas Nova
%    \and
%    Gaëlle Molinari
%}
%\institute{
%    Pierre Dillenbourg, Séverin Lemaignan \at Computer-Human Interaction in Learning and Instruction \\
%    École Polytechnique Fédérale de Lausanne (EPFL) \\ CH-1015 Lausanne, Switzerland \\
%    \email{firstname.lastname@epfl.ch}
%    \and
%    Nicolas Nova \at Haute École Arts \& Design de Genève \\ CH-1201 Genève, Switzerland\\\email{nicolas.nova@hesge.ch}
%    \and
%    Gaëlle Molinari \at Distance Learning University Switzerland (Unidistance),
%    CH-3960 Sierre, Switzerland\\\email{gaelle.molinari@unidistance.ch}
%}
\author{}
\begin{document}
\maketitle

@@

From its very beginning, CSCL research has been following
\citet{roschelle1995construction} suggestion that collaborative learning has
something to do with the process of constructing and maintaining a \emph{shared
understanding} of the task at hand. Building a shared/mutual understanding
refers to the upper class of collaborative learning situations, those in which
students should build upon each other's understanding to refine their own
understanding: what is expected to produce learning is not the mere fact that
two students build the same understanding but the cognitive effort they have to
engage to build this shared understanding~\citep{schwartz1995emergence}. This
effort can be observed by the frequency of rich interactions, \ie interactions
whose occurrence has been related to learning: (self-) explanations in cognitive
science~\citep{chi1989self, webb1991task}, conflict resolution in
socio-cognitive theories~\citep{doise1975social},
mutual regulation~\citep{blaye1995collaborative} in a Vygostkian perspective, etc. The
construction of a shared understanding has been investigated for several years
in psycholinguistics, under the  notion of
\emph{grounding}~\citep{clark1986referring}. However, the relevance of grounding
mechanisms for explaining learning outcomes has been questioned in learning
sciences. The monitoring and repair of misunderstanding explains for instance
referential failures in short dialogue episodes but does hardly predict
\emph{conceptual change} (\ie the acquisition, acceptation and integration of a
new belief into one's mental model) over longer
sessions~\citep{dillenbourg2006sharing}. The cumulative effect of grounding
episodes can probably be better understood from a socio-cultural perspective:

@@

This interesting cognitive science debate occurred mostly outside the field of
learning. In education, the question is to relate these mechanisms to learning
outcomes: \textbf{Is linguistic alignment sufficient to trigger conceptual
change?} \textbf{Does negotiation of meaning only occurs when partners monitor
and diagnose each other's knowledge?} If the ratio between shallow alignment and
deep grounding depends upon the task, and if deep grounding is a condition for
learning, then the pedagogical challenge is indeed to design tasks that require
deep grounding. Most empirical studies on grounding and alignment are conducted
with tasks, which despite being qualified as ``ecologically valid'' by their
authors, are mere referencing tasks such as asking the way to the train station
or helping the peer to choose a picture among many. In this contribution, we
explore several richer tasks such as arguing on a sensitive issue or building a
concept map.

@@

The content of mutual models ranges from \emph{dispositional} versus
\emph{situational} aspects. The \emph{dispositional} aspects refer to $A$'s
representation of $B$'s long term knowledge, skills or traits. It is thus closely
related to the notion of transactive memory~\citep{wegner1987transactive,
moreland1999transactive}.  \emph{Situational} aspects refer to $A$'s representation of
B's knowledge, behavior or intentions specifically activated in the situation in
which $A$ and $B$ are collaborating, some of them being valid for 2 seconds, other
ones for 2 hours. Examples of fragments that constitute $A$'s model of B
regarding to aspects X, \ie $Model(A,B,X)$, abbreviated \model{A}{B}{X}, could be:

@@

    \item $cost(\alpha)$: in all studies but study 4, the cost of  grounding is
        an independent variable. Study 3 uses media richness as independent
        variable, with the hypothesis that modelling emotions is ``cheaper'' with a
        richer medium, \ie  when peers can see each other.  Studies 1,2 and 4
        use awareness tools which, in principle, reduce the cost of mutual modelling, but do
        not eliminate all costs: if the tool provide $A$ with information about
        what $B$ does/knows, this additional information may actually increase
        cognitive load. Awareness tools constitute a kind of mutual modelling prosthesis,
        and, like any prosthesis, they may augment mutual modelling (by facilitating it or
        even scaffolding it) or inhibit it (by making it useless).

@@

{\sc SpaceMiners} is a 3D game that involves two players collecting minerals located
in asteroids (Fig.~\ref{study1:spaceminer}). To do so, they must launch drones through
the space after choosing their initial direction and speed. Once launched, the
trajectory of drones is only influenced by the gravity of planets and by
trajectory modification tools.  During the experiment, the teams were confronted
with three increasingly complex situations. The experiment was 2 hours long: a
30 minutes tutorial and 3 levels of 30 minutes. Players uses using a regular
joystick and communicated with each other through an audio channel.

@@

The results evidence that richer media does not automatically foster better
modelling, on the contrary: groups using only the audio modality to exchange
their views were able to build more accurate models of each others than groups
using both video and audio, which is especially surprising when modelling
emotions: one would expect that video feedback indeed help building a better
model of the emotions of the other. Our interpretation is that viewing what
one's partner sees (shared graphical editor) was, in this task, more important
than seeing each other~\citep{gaver1993one,anderson1997impact}.

>>>>

% ANON
%\author{
%    Pierre Dillenbourg
%    \and
%    Séverin Lemaignan
%    \and
%    Mirweis Sangin
%    \and
%    Nicolas Nova
%    \and
%    Gaëlle Molinari
%}
%\institute{
%    Pierre Dillenbourg, Séverin Lemaignan \at Computer-Human Interaction in Learning and Instruction \\
%    École Polytechnique Fédérale de Lausanne (EPFL) \\ CH-1015 Lausanne, Switzerland \\
%    \email{firstname.lastname@epfl.ch}
%    \and
%    Mirweis Sangin \at Use-able consulting \\ CH-1004 Lausanne, Switzerland\\\email{mirweis@gmail.com}
%    \and
%    Nicolas Nova \at Haute École Arts \& Design de Genève \\ CH-1201 Genève, Switzerland\\\email{nicolas.nova@hesge.ch}
%    \and
%    Gaëlle Molinari \at Distance Learning University Switzerland (Unidistance),
%    CH-3960 Sierre, Switzerland\\\email{gaelle.molinari@unidistance.ch}
%}
\author{}
\begin{document}
\maketitle

@@

From its very beginning, CSCL research has been following
the suggestion by \citet{roschelle1995construction} that collaborative learning has
something to do with the process of constructing and maintaining a \emph{shared
understanding} of the task at hand. Building a shared/mutual understanding
refers to the upper class of collaborative learning situations, those in which
students should build upon each other's understanding to refine their own
understanding: what is expected to produce learning is not the mere fact that
two students build the same understanding but the cognitive effort they have to
engage to build this shared understanding~\citep{schwartz1995emergence}. This
effort can be observed by the frequency of rich interactions, \ie interactions
whose occurrence has been related to learning: (self-) explanations in cognitive
science~\citep{chi1989self, webb1991task}, conflict resolution in
socio-cognitive theories~\citep{doise1975social},
mutual regulation~\citep{blaye1995collaborative} in a Vygostkian perspective, etc. The
construction of a shared understanding has been investigated for several decades
in psycholinguistics, under the  notion of
\emph{grounding}~\citep{clark1986referring}. However, the relevance of grounding
mechanisms for explaining learning outcomes has been questioned in the learning
sciences. The monitoring and repair of misunderstanding explains for instance
referential failures in short dialogue episodes but does hardly predict
\emph{conceptual change} (\ie the acquisition, acceptation and integration of a
new belief into one's mental model) over longer
sessions~\citep{dillenbourg2006sharing}. The cumulative effect of grounding
episodes can probably be better understood from a socio-cultural perspective:

@@

This interesting cognitive science debate occurred mostly outside the field of
learning. In education, the question is to relate these mechanisms to learning
outcomes: \textbf{Is linguistic alignment sufficient to trigger conceptual
change?} \textbf{Does negotiation of meaning only occur when partners monitor
and diagnose each other's knowledge?} If the ratio between shallow alignment and
deep grounding depends upon the task, and if deep grounding is a condition for
learning, then the pedagogical challenge is indeed to design tasks that require
deep grounding. Most empirical studies on grounding and alignment are conducted
with tasks that, despite being qualified as ``ecologically valid'' by their
authors, are mere referencing tasks such as asking the way to the train station
or helping the peer to choose a picture among many. In this contribution, we
explore several richer tasks such as arguing on a sensitive issue or building a
concept map.

@@

The content of mutual models ranges from \emph{dispositional} to
\emph{situational} aspects. The \emph{dispositional} aspects refer to $A$'s
representation of $B$'s long term knowledge, skills or traits. It is thus closely
related to the notion of transactive memory~\citep{wegner1987transactive,
moreland1999transactive}.  \emph{Situational} aspects refer to $A$'s representation of
B's knowledge, behavior or intentions specifically activated in the situation in
which $A$ and $B$ are collaborating, some of them being valid for 2 seconds, other
ones for 2 hours. Examples of fragments that constitute $A$'s model of B
regarding to aspects X, \ie $Model(A,B,X)$, abbreviated \model{A}{B}{X}, could be:

@@

    \item $cost(\alpha)$: in all studies but study 4, the cost of  grounding is
        an independent variable. Study 3 uses media richness as independent
        variable, with the hypothesis that modelling emotions is ``cheaper'' with a
        richer medium, \ie  when peers can see each other.  Studies 1,2 and 4
        use awareness tools which, in principle, reduce the cost of mutual modelling, but do
        not eliminate all costs: if the tool provides $A$ with information about
        what $B$ does/knows, this additional information may actually increase
        cognitive load. Awareness tools constitute a kind of mutual modelling prosthesis,
        and, like any prosthesis, they may augment mutual modelling (by facilitating it or
        even scaffolding it) or inhibit it (by making it useless).

@@

{\sc SpaceMiners} is a 3D game that involves two players collecting minerals located
in asteroids (Fig.~\ref{study1:spaceminer}). To do so, they must launch drones through
the space after choosing their initial direction and speed. Once launched, the
trajectory of drones is only influenced by the gravity of planets and by
trajectory modification tools.  During the experiment, the teams were confronted
with three increasingly complex situations. The experiment was 2 hours long: a
30 minute tutorial and 3 levels of 30 minutes. Players uses using a regular
joystick and communicated with each other through an audio channel.

@@

The results evidence that richer media does not automatically foster better
modelling, on the contrary: groups using only the audio modality to exchange
their views were able to build more accurate models of each other than groups
using both video and audio, which is especially surprising when modelling
emotions: one would expect that video feedback indeed help building a better
model of the emotions of the other. Our interpretation is that viewing what
one's partner sees (shared graphical editor) was, in this task, more important
than seeing each other~\citep{gaver1993one,anderson1997impact}.
